Verification of Stencil Computations through Spatial Specifications

### 2016-07-05 paper.tex

@@

\title{Verification of Stencil Computations through Spatial Specifications}
\authorinfo{Dominic Orchard \and Mistral Contrastin
\and Matthew Danish \and Andrew Rice}{University of Cambridge}{}

@@

\begin{abstract}
  Verifying the correctness of numerical computations relies on first
  giving a specification of their behaviour. In many cases, a
  numerical computation is derived from an underlying mathematical
  model, but this is often difficult to use as a specification as it
  is highly decoupled from the implementation. We have been studying
  specification techniques in middle ground between low-level and
  high-level specifications for numerical code. In this paper, we
  focus on \emph{stencil computations}---a common idiom in numerical code, but
  one which is error-prone due to fine-grained indexing errors. We
  hypothesise that in practise, stencil computations tend to have a
  regular shape which is amenable to static analysis and simple
  specification. We describe a abstract specification language for
  stencil computations that can be directly embedded as annotations in
  source code, along with an inference and checking procedure. We
  evaluate our language against a corpus of numerical Fortran code,
  and show that precise specifications can be given to roughly half of
  the identifiable stencil computations, and that the vast majority of
  stencil computations indeed have a simple, regular, static
  shape. This paper details our stencil specification language, its
  static semantics, inference, checking, and program synthesis
  procedures, our implementation, and our evaluation study. We provide
  an implementation of stencil inference and validation in Camfort, an
  open-source program analysis and refactoring tool.
\end{abstract}

@@

\begin{definition}The product of two models of dimensionality
$n$ is given by the tensor $\otimes_n$ defined:
%%
\newcommand{\effdims}[2]{\mathit{dims}(#1)_{#2}}
\begin{alignat*}{2}
& \effdims{M}{n} =
\bigcup_{u \in M} \{x \mid x \in \{1,...,n\}, u_x \neq \infty\} \\
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{5.6cm}{$i \in \{1, \ldots, 2^n\},$
                   $(u, v) \in M \times N,$ \\
                   $x = (u \bowtie v)_i,$ \\
                   $j \in (\effdims{M}{n} \cup \effdims{N}{n}),$ \\
                   $x_j \neq \infty$
                  } \Bigg\}
\end{alignat*}
The $\bowtie$ operation, which we call \emph{pairwise permutation}
(described below), constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors). Using this, the $\otimes$ operation constructs the set of every
pairwise permutation for every pair $(u, v)$ of $n$-vectors in the Cartesian
product of $M \times N$.

@@

\onecolumn
\appendix

\input{proof.tex}

>>>>

%% MODEL
\newcommand{\effdims}[2]{\mathit{dims}(#1)_{#2}}

@@

\title{Verification of Stencil Computations through Spatial
  Specifications}
\authorinfo{Blind}{}{}
%\authorinfo{Dominic Orchard \and Mistral Contrastin
%\and Matthew Danish \and Andrew Rice}{University of Cambridge}{}

@@

\begin{abstract}
  Verifying the correctness of numerical computations relies on first
  giving a specification of their behaviour. In many cases, a
  numerical computation is derived from an underlying mathematical
  model, but this is often difficult to use as a specification as it
  is highly decoupled from the implementation. We have been studying
  specification techniques in middle ground between low-level and
  high-level specifications for numerical code. In this paper, we
  focus on \emph{stencil computations}---a common idiom in numerical code, but
  one which is error-prone due to fine-grained indexing errors. We
  hypothesise that in practise, stencil computations tend to have a
  regular shape which is amenable to static analysis and simple
  specification. We describe a abstract specification language for
  stencil computations that can be directly embedded as annotations in
  source code, along with an inference and checking procedure. We
  evaluate our language against a corpus of numerical Fortran code,
  and show that precise specifications can be given to roughly half of
  the identifiable stencil computations, and that the vast majority of
  stencil computations indeed have a simple, regular, static
  shape. This paper details our stencil specification language, its
  static semantics, inference, checking, and program synthesis
  procedures, our implementation, and our evaluation study. We provide
  an implementation of ontop of CamFort, an
  open-source program analysis and refactoring tool for Fortran corpus.
\end{abstract}

@@

\begin{definition}The product of two models of dimensionality
$n$ is given by the tensor $\otimes_n$ defined:
%%
\begin{alignat*}{2}
& \effdims{M}{n} =
\bigcup_{u \in M} \{x \mid x \in \{1,...,n\}, u_x \neq \infty\} \\
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{5.6cm}{$i \in \{1, \ldots, 2^n\},$
                   $(u, v) \in M \times N,$ \\
                   $x = (u \bowtie v)_i,$ \\
                   $j \in (\effdims{M}{n} \cup \effdims{N}{n}),$ \\
                   $x_j \neq \infty$
                  } \Bigg\}
\end{alignat*}
The $\bowtie$ operation, which we call \emph{pairwise permutation}
(described below), constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors). Using this, the $\otimes$ operation constructs the set of every
pairwise permutation for every pair $(u, v)$ of $n$-vectors in the Cartesian
product of $M \times N$.

@@
